Supplementary Materials

### 2020-05-05 analysis/analysis.Rmd
Sup mat new version up to PPC

---
title: "Bayesian Modelling"
author: "Alasdair DF Clarke"
date: "21 October 2019"
output: 
  html_document:
    number_sections: true
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)

# specify some colours to use in the plots
v_cols <- ggthemes::ptol_pal()(6)
```

@@

```{r packages, results='hold'}
library(tidyverse)
library(brms)
library(tidybayes)
library(ggridges)
library(latex2exp)
library(see)
library(ggthemes)
```

@@

Import data and quickly plot.

@@

Importing the eeg occ odd data.

```{r eeg_import, cache=TRUE}
fl <- "rms_Occ_Odd.csv"

@@

```{r}
d_eeg %>% ggplot(aes(x = rms)) + 
    geom_histogram(binwidth = 0.2, fill = "skyblue", colour = "black") +
    theme_lucid()
```

@@

```{r threshold_import, cache=TRUE}
fl <- "thresholds"

@@

read_csv("../Experiment/analysis/thresholds.csv", col_types = my_cols_threshold) %>%
  select(subject = "person", wallpaper_group = "group", threshold) %>%
  glimpse() -> d_dispthresh

@@

As above, eyeball the data…

```{r}
d_dispthresh %>% ggplot(aes(x = threshold)) + 
    geom_histogram(binwidth = 0.2, fill = "skyblue", colour = "black") +
    theme_lucid()
```
Again, we have a skewed distribution, so will fit with `family = 'lognormal'`.

@@

We will also include some control EEG data. These are the …. < details from Peter >

@@

*  Use weakly-informative priors.
*  Fit a multi-level model using the lognormal distribution.
  + independent variable (wallpaper group) and one dependent variable (rms/threshold).
  +maximal random effect structure.
* Check model fit (r̂ statistics and trace plots).
* Extract mcmc samples from the posterior distributions.
* Use these samples to estimate the difference between sub- and super-grounps.

@@

First we have to define some priors: are we happy with these?

We will keep the default priors for the group-level (‘random’) effects. From the brms documentation:

@@

Now we can fit a Bayesian multi-level model. First we will run with sample_prior = 'only' to check that our priors appear reasonable.

```{r}
## define some priors
prior <- c(
  set_prior("normal(0,2)", class = "b"),    
  set_prior("cauchy(0,2)", class = "sigma"))
```

Now we can fit a Bayesian multi-level model. First we will run with `sample_prior = 'only'` to check that our priors appear reasonable.

```{r eeg_prior_pred_check, cache=TRUE}
m_prior <- brm(data = d_eeg, 
  rms ~ wallpaper_group-1 + (1|subject),
  family = "lognormal", 
  prior = prior, 
  iter = n_iter ,
  sample_prior = 'only')
```

Now we will define some handy functions so we can reuse the same code to visualise the results of the model

```{r}
get_model_samples <- function(my_model) {

  my_samples <- posterior_samples(my_model)

  names(my_samples)  %>% 
    gsub('[b_]*wallpaper_group', "", .) -> names(my_samples)

  return(my_samples)
}

get_model_predictions <- function(my_samples) {
  my_samples %>%
    as_tibble() %>% 
    mutate_all(exp) %>%
    select(P2, PM, PG, CM, PMM, PMG, PGG, CMM, P4, P4M, P4G, P3, P3M1, P31M, P6, P6M) %>%
    gather() -> my_model_predictions

  return(my_model_predictions)
}

plot_model_output <- function(mdat, edat, my_title, my_colour, x_label) {
  x_var <- names(edat)[3]
  # plot model fit (mdat) against empiricaly data (edat)
  ggplot(mdat, aes(x = value, y = key)) +
      stat_pointintervalh(.width = c(.66, .95), colour = my_colour) +
      geom_density_ridges(data = edat,aes_string(x = x_var, "wallpaper_group"),
        fill = "grey",
        alpha = 0.25,
        scale = 2,
        bandwidth = 0.05) +
      scale_x_log10(x_label) +
    ggtitle(my_title) +
      theme_lucid()
}
```

Now we can use these functions to visualise the predictions made by the model:

```{r}

prior_samples <- get_model_samples(m_prior) %>% glimpse()

@@

# # And now plot
plot_model_output(prior_predictions, d_eeg,
                   "Prior Predictive Check: EEG RMS", "deepskyblue3", "rms")

```

The density plots show the distribution of the data, while the point interval shows the median with $66\%$ and $95\%$ quantiles for our prior.

It looks like our priors are in the right range for the data.

### Prior Predictive Check for Display Duration Data

Now we can repeat this process (with the same priors) for the display duration threshold data.

```{r dispthresh_prior_pred_check, cache=TRUE, echo=FALSE}
m_prior <- brm(data = d_dispthresh, 
               threshold ~ wallpaper_group-1 + (1|subject),
               family = "lognormal", 
               prior = prior, iter = n_iter ,
               sample_prior = 'only')

# Extract samples from model
prior_samples <- get_model_samples(m_prior)

# Tidy up a little
prior_predictions <- get_model_predictions(prior_samples)

plot_model_output(prior_predictions, d_dispthresh, 
                  "Prior Predictive Check: Display Duration Thresholds", "deepskyblue3", "display threshold (ms)")
```

The density plots show the distribution of the data, while the point interval shows the median with $66\%$ and $95\%$ quantiles for our prior.

It looks like our priors are in the right range for the data.

## Compute Posterior

### Fit Model to EEG Data

We will now fit the model to the data.

```{r eeg_fit_model, cache=TRUE}
m_eeg <- brm(data = d_eeg, 
             rms ~ wallpaper_group-1 + (1|subject),
             family = "lognormal", 
             prior = prior, iter = n_iter,
             control = list(adapt_delta = 0.99))

summary(m_eeg)
```

Hopefully all the $\hat{r} = 1$ and there aren't any warning messages, etc.

We will now have at what the model predicts for the average participant (i.e, ignoring the random intercepts). If anybody else would like to suggest over sensible ways to summarise this model, do let me know! I've been trying to use some of the functions from the `tidybayes` package, and they don't (yet?) have much on random effects.

```{r}
 # Extract samples from model

my_samples$eeg <- get_model_samples(m_eeg)

 # Tidy up a little
 my_models$eeg <- get_model_predictions(my_samples$eeg)

 # Plot
 plot_model_output(my_models$eeg, d_eeg, 
                   "Model Fits: EEG RMS", "orangered3", "rms")
```

### Fit Model to Display Duration Threshold Data

We will now fit the model to the data.

```{r threshold_fit_model, cache=TRUE, echo=FALSE}
m_threshold <- brm(data = d_dispthresh, 
                   threshold ~ wallpaper_group-1 + (1|subject),
                   family = "lognormal", 
                   prior = prior, iter = n_iter,
                   control = list(adapt_delta = 0.99))

summary(m_threshold)
```

Hopefully all the $\hat{r} = 1$ and there aren't any warning messages, etc.

We will now have at what the model predicts for the average participant (i.e, ignoring the random intercepts). If anybody else would like to suggest over sensible ways to summarise this model, do let me know! I've been trying to use some of the functions from the `tidybayes` package, and they don't (yet?) have much on random effects.

```{r}
# Extract samples from model
my_samples$threshold <- get_model_samples(m_threshold)

# Tidy up a little
my_models$threshold <- get_model_predictions(my_samples$threshold)

# Plot
plot_model_output(my_models$threshold, d_dispthresh, 
                  "Model Fits: Display Duration Threshold", "orangered3", "threshold (ms)")
```

### EEG Control Data

We will also fit models to the blah and blah data as control conditions.

First for the occ_even data...

```{r occ_even_model, cache=TRUE}

m_eeg_occ_even <- brm(data = d_eeg_occ_even, 
                      rms ~ wallpaper_group-1 + (1|subject),
                      family = "lognormal", 
                      prior = prior, iter = n_iter,
                      control = list(adapt_delta = 0.99))

# Extract samples from model
my_samples$occ_even <- get_model_samples(m_eeg_occ_even)
```

And now for th par_odd data...

```{r par_odd_model, cache=TRUE}

m_eeg_par_odd <- brm(data = d_eeg_par_odd, 
                     rms ~ wallpaper_group-1 + (1|subject),
                     family = "lognormal", 
                     prior = prior, iter = n_iter,
                     control = list(adapt_delta = 0.99))

my_samples$par_odd <- get_model_samples(m_eeg_par_odd)
```

# Subgroup Comparisons

Helper function to do make comparisons and create plot:

```{r}
source("get_subgroup_comparisons.R")
```

```{r subgroup_comparisons, cache=TRUE}

subgroup_comp$eeg <- get_subgroup_comparisons(my_samples$eeg, 2*n_iter) %>% glimpse()
subgroup_comp$threshold <- get_subgroup_comparisons(my_samples$threshold, 2*n_iter)

# Also compute for control datasets
subgroup_comp$occ_even <- get_subgroup_comparisons(my_samples$occ_even, 2*n_iter) 
subgroup_comp$par_odd  <- get_subgroup_comparisons(my_samples$par_odd,  2*n_iter)

```

## EEG RMS

Finally, I calculate the probability that the difference between subgroup and supergroup is larger than zero (small than zero for the display duration) given the data. This information is then binned so we can colour in the posterior density plots.

First we will define a handy function that we can reuse a few times.

```{r define_comparison_summary_function, echo=TRUE}
comparison_summary <- function(subgroup_comp_dat, greater, measurement_lab) {
  # compute probabilties
  subgroup_comp_dat %>%
    group_by(key, index, normal) %>%
    summarise(
      mean_value = mean(value),
      prob_one_tailed = if_else(greater == TRUE, mean(value > 0), mean(value < 0)))  %>%
    mutate(
      measurement = measurement_lab,
      index_labs = as.numeric(as.factor(index)),
      lab_cols = v_cols[index_labs],
      prob_cat = cut(prob_one_tailed, c(0, 0.5, 0.75, 0.9, 0.99, 1))) -> comp_summary_dat

  return(comp_summary_dat)
}
```

Now we will use it!

```{r, plot_eeg_comp, fig.height=8}
# compute probabilties
comp_summary$eeg <- comparison_summary(subgroup_comp$eeg, TRUE, 'rms') %>% glimpse()

subgroup_comp$eeg <- full_join(
  subgroup_comp$eeg, 
  comp_summary$eeg, 
  by = c("key", "index", "normal"))

# Finally, plot everything:
p1 <- plot_comparisons(
  filter(subgroup_comp$eeg, index == 2),
  filter(comp_summary$eeg, index == 2), 
  1)

p2 <-plot_comparisons(
  filter(subgroup_comp$eeg, index != 2),
  filter(comp_summary$eeg, index != 2), 
  2)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
both_plts
```

## Display Duration Threshold

We can do the same for the display duration thresholds.

```{r plot_threshold_comp, fig.height=8}

# compute probabilties
comp_summary$threshold <- comparison_summary(subgroup_comp$threshold, FALSE, 'threshold') %>% glimpse()

subgroup_comp$threshold <- full_join(
  subgroup_comp$threshold, 
  comp_summary$threshold, 
  by = c("key", "index", "normal"))

# Finally, plot everything:
p1 <- plot_comparisons(
  filter(subgroup_comp$threshold, index == 2),
  filter(comp_summary$threshold, index == 2), 
  1, TRUE)
p2 <- plot_comparisons(
  filter(subgroup_comp$threshold, index != 2),
  filter(comp_summary$threshold, index != 2), 
  2, TRUE)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
both_plts
```

## Control Data

We will now do exactly the same with the control data...

```{r comp_control, fig.height=8, echo = FALSE}

# compute probabilties
comp_summary$occ_even <- comparison_summary(subgroup_comp$occ_even, TRUE, 'rms')

subgroup_comp$occ_even <- full_join(
  subgroup_comp$occ_even, 
  comp_summary$occ_even, 
  by = c("key", "index", "normal"))

p1 <- plot_comparisons(
  filter(subgroup_comp$occ_even, index == 2),
  filter(comp_summary$occ_even, index == 2), 
  1)
p2 <- plot_comparisons(
  filter(subgroup_comp$occ_even, index != 2),
  filter(comp_summary$occ_even, index != 2), 
  2)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv") 
both_plts

# compute probabilties
comp_summary$par_odd <- comparison_summary(subgroup_comp$par_odd, TRUE, 'rms')

subgroup_comp$par_odd <- full_join(
  subgroup_comp$par_odd, 
  comp_summary$par_odd, 
  by = c("key", "index", "normal"))

p1 <- plot_comparisons(
  filter(subgroup_comp$par_odd, index == 2),
  filter(comp_summary$par_odd, index == 2), 
  1)
p2 <- plot_comparisons(
  filter(subgroup_comp$par_odd, index != 2),
  filter(comp_summary$par_odd, index != 2), 
  2)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
both_plts

```

## Summary

Some summary stats for how many subgroups show a difference. We have 64 subgroupds in total.

```{r}

count_comparisons <- function(comparisons, p) {
  x <- sum(comparisons$prob_one_tailed >= p)
  return(x)
}

df <- tibble()

for (p in seq(0.5, 0.99, 0.01)) {

  df <- bind_rows(df, map_dbl(comp_summary,  count_comparisons, p))
}

df$p <- seq(0.5, 0.99, 0.01)

df %>% gather("measurement", "number", -p) %>%
  mutate(proportion = number/64) %>%
  ggplot(aes(x = p, y = proportion, colour = measurement)) + geom_point() + geom_path() +
  theme_minimal() + ggthemes::scale_color_ptol() +
  xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")

filter(df, p == 0.95)

```

# Additional Analysis

## Index and Normality

## Correlation Between EEG RMS and Display Duration Threshold

```{r compare_comparisons}

bind_rows(
  select(comp_summary$eeg, measurement, key, mean_value),
  select(comp_summary$threshold, measurement, key, mean_value)) %>%
  spread(measurement, mean_value) %>%
  glimpse()-> d

m <- brm(data = d, threshold ~ rms)

summary(m)

bayes_R2(m)

d %>% 
  modelr::data_grid(rms = seq(-0.2, 0.8, 0.1)) %>%
  add_fitted_draws(m, n = 100) %>%
  ggplot(aes(x = rms, y = threshold, color = index)) +
  geom_line(aes(y = .value, group = .draw), alpha = .1, colour = "thistle4") +
  geom_point(data = d) +
  theme_minimal() + 
  ggthemes::scale_colour_ptol()

```

>>>>

---
title: "Supplementary Materials"
author: "Peter J. Kohler, Oda Åsgard, and Alasdair D. F. Clarke"
date:
output: bookdown::html_document2
---

@@

```{r setup, include=FALSE}

library(bookdown)
library(tidyverse)
library(brms)
library(tidybayes)
library(ggridges)
library(latex2exp)
library(see)
library(ggthemes)
library(patchwork)

knitr::opts_chunk$set(echo = FALSE, message=FALSE, fig.align='center')

# specify some colours to use in the plots
v_cols <- ggthemes::ptol_pal()(6)

# read in some helper functions that will be used later
source("scripts/helper_functions.R")
```

@@

@@

Here we import our data and make some sumamry plots.

@@

Importing the odd-harmonic EEG data from six electrodes region-of-interest over occipital cortex.

```{r eeg_import}
fl <- "rms_Occ_Odd.csv"

@@

```{r cache=TRUE, fig.cap = "Data are the root-mean-squaured (rms) over the odd-harmonic filtered waveforms."}
d_eeg %>% ggplot(aes(x = rms)) + 
    geom_histogram(binwidth = 0.2, fill = "skyblue", colour = "black") +
    theme_lucid()
```

@@

```{r threshold_import}
fl <- "thresholds"

@@

read_csv("../Experiment/process_results/thresholds.csv", col_types = my_cols_threshold) %>%
  select(subject = "person", wallpaper_group = "group", threshold) %>%
  glimpse() -> d_dispthresh

@@

As above, a summary of the data.

```{r, fig.cap="Histogram of the display duration thresholds."}
d_dispthresh %>% ggplot(aes(x = threshold)) + 
    geom_histogram(binwidth = 0.2, fill = "skyblue", colour = "black") +
    theme_lucid()
```

Again, we have a skewed distribution, so will fit with `family = 'lognormal'`.

@@

In addition to the primary EEG data set, we are also importing two control data sets which are (a) even harmonic data from the same occipital electrodes, and (b) odd harmonic data 
from six parietal electrodes (see main paper).

@@

*  Use weakly-informative priors.
*  Fit a multi-level model using the lognormal distribution.
  + independent variable (wallpaper group) and one dependent variable (rms or threshold).
  + maximal random effect structure.
* Check model fit (r̂ statistics and trace plots).
* Extract mcmc samples from the posterior distributions.
* Use these samples to estimate the difference between sub- and super-groups.

@@

In this section we will specify some priors. We then then use a prior-predictive check to assess whether the prior is reasonable or not (i.e., on the same order of magnitude as our measurements).

### Fixed Effects

Our independent variable is a categorical factor with 16 levels. We will drop the intercept from our model and instead fit a coefficent for each factor level ($y \sim x - 0$). As our dependant variable will be log-transformed, we can use the priors below:

```{r, echo = TRUE}
prior <- c(
  set_prior("normal(0,2)", class = "b"),    
  set_prior("cauchy(0,2)", class = "sigma"))
```

### Group-level Effects

We will keep the default weakly informative priors for the group-level (‘random’) effects. From the brms documentation:

@@

### Prior Predictive Check

Now we can specify our Bayesian multi-level model and priors. Note that as we are using  `sample_prior = 'only'`, the model will not learn anything from our data.

```{r eeg_prior_pred_check, cache=TRUE, echo=TRUE}
m_prior <- brm(data = d_eeg, 
  rms ~ wallpaper_group-1 + (1|subject),
  family = "lognormal", 
  prior = prior, 
  iter = n_iter ,
  sample_prior = 'only')
```

We can use this model to generate data.

```{r, fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
prior_samples <- get_model_samples(m_prior)

@@

# # And now plot
p1 <- plot_model_output(prior_predictions, d_eeg,
                   "odd-harmonic data from \noccipital electrodes", "cornflowerblue", "log(rms)")

p2 <- plot_model_output(prior_predictions, d_dispthresh,
                   "display duration threshold data", "cornflowerblue", "log(ms)")

# draw pot
plt <- p1 + p2
plt + plot_annotation(title = 'Prior Predictive Checks')
```

We can see that i) our priors are relatively weak as the predictions span several orders of magnitide, and ii) our empirical data falls within this range.
